Work Fragmentation in Developer Interaction Data

### 2015-08-07 paper.tex
section 8

On refactoring activities, Murphy-Hll et al. used this data to understand how programmers are using refactoring tools analyzing patterns on refactoring practices, finding that most programmers don't make deep use of these tools, leaving untouched most of the configuration parameters and performing manually most of the refactoring  \cite{MPB12}. Liu et al. identified refactoring tactics (floss refactoring and root canal refactoring, following Murphy-Hill's terminology), finding that the majority of them are floss refactoring and detecting differences on activities performed depending on the type of the refactoring \cite{LGN12}. Sahin et al. analyzed the energy usage changes after refactoring and used the UDC data to identify the most common refactoring events and selected those that fulfill the requirements of being common and cause structural changes \cite{SPC14}.

@@

\section{Data Description}

@@

\section{Usage Data Collector}

\subsection{Data description}
The Usage Data Collector dataset is a big compendium of information about interaction data from users of Eclipse through December 2008 to August 2010 with the intention to keep record of how programmers are using the IDE. It's a framework that listens to the events triggered by the user or system like edition and navigation commands, the startup of a plug in or the closing of the platform. More specifically, UDC collects information about loaded bundles, commands accessed via keyboard shortcuts, actions invoked via menu or tool-bars, perspective changes, view usage and editor usage. It is a big dataset that contains information of around 1,800,000 users and has a total of 2,323,233,101 unique rows with 5 attributes each. The table \ref{tbl:att_udc} show a description of the attributes.

@@

\subsection{Selection and transformation}
Due to the magnitude of the dataset we only worked on a fragment of it. By doing a random selection of 1,000 users we obtained a good representation of it all, which is actually bigger and richer that the dataset we had from Mylyn.

We obtained this version of the data from a project published on Google BigQuery by Emerson Murphy-Hill [ref]. It is the same data but cleaned and preprocessed so the transformation phase is simple and focused on our needs.

@@

After obtaining the data, the next step was to add the attributes required by every event to match it to the data on Mylyn, such is the type of the event, the duration and an ID to identify the different sessions present on the data. For the latter, we sorted the data by \textit{userId} and by \textit{datetime} afterwards. This was required because by default the user's data is mixed and we needed it not only chronologically correct but also sorted by users to tag the sessions of every user without interferences.

With Mylyn we have information about the end time of an event. The UDC dataset doesn't have an attribute to show this information, so we added an interval time in seconds that depends on the next event executed. This is important to identify those events that represent an interruption or the end of a session.

Once the events were classified and sorted, we split the dataset into sessions, defining a session as a period of registered work between interruptions of 8 hours or larger, as we did with Mylyn. We did this splitting the data by users to avoid a conflict between the last event of a user and the first one of the next. After this, we obtained a total of 28,989 different sessions, but after a filter to keep only those with at least 30 minutes of productive time we ended up with 15,825 useful sessions, in contrast with 4,284 we had from Mylyn.

\subsection{Classification of events}
To select the type, we made a thorough analysis of the different events to classify them where they better fit. In comparison with Mylyn we didn't have a classification by default but the attributes can give us a very good approximation.

We took into consideration the attributes \textit{what}, \textit{kind} and \textit{description} to choose the best classification for every event. Starting by the kind attribute, some events are labeled as \textit{view} which are precisely those that open, activate or close a view such as the console, the package explorer or the variables view. Therefore, the events of the kind \textit{view} were classified as selection.

The classification for the rest of the events is more complex because we needed to check directly the description and bundle. We could label as edition all those commands with the attributes \textit{command} and \textit{executed}, and the majority of them are in fact of that type, but some commands executed also can open a view, navigate through the text or go to another class. For this reason we took a closer look into the bundle and what the description can tell us about the events.

By default, all events of the attributes \textit{command} and \textit{executed} were classified as edition and afterwards we did a thoroughly analysis to identify those that are actually selection events. This way we categorized as selection all the events that we can tell on the description if they have the purpose to navigate to another window, open a perspective, find and replace words, search, window management and debug. The description contains the whole path from the top package to the class that implements the event, so we look for key words to apply the classification. The proportion of events after the classification is shown in the table \ref{tbl:prop_events_udc}.

@@

The edition and selection types are of the same nature as with Mylyn. The system classification is composed of the events that activate and deactivate the workspace of Eclipse; It is difficult to determine when and why this events are executed (they can be executed either via user or system action, and the order of execution in relation with associated events is not clear) so we decided to establish this special classification for them. Finally, the events of the type control are those for the control version of files.

>>>>

On refactoring activities, Murphy-Hill et al. used this data to understand how programmers are using refactoring tools analyzing patterns on refactoring practices, finding that most programmers don't make deep use of these tools, leaving untouched most of the configuration parameters and performing manually most of the refactoring  \cite{MPB12}. Liu et al. identified refactoring tactics (floss refactoring and root canal refactoring, following Murphy-Hill's terminology), finding that the majority of them are floss refactoring and detecting differences on activities performed depending on the type of the refactoring \cite{LGN12}. Sahin et al. analyzed the energy usage changes after refactoring and used the UDC data to identify the most common refactoring events and selected those that fulfill the requirements of being common and cause structural changes \cite{SPC14}.

@@

\section{Mylyn Data Description}

@@

\section{Eclipse Usage Data Collector Data Description}

\subsection{The Eclipse UDC}
The Usage Data Collector (UDC) dataset is a large compendium of information about interaction data from users of Eclipse, collected from December 2008 to August 2010, with the intention to keep track of how programmers are using the IDE. The framework listens to the events triggered by the user or the system, such as: edition and navigation commands; the startup of a plug in; or the closing of the platform. More specifically, UDC collects information about loaded bundles, commands accessed via keyboard shortcuts, actions invoked via menu or tool-bars, perspective changes, view usage and editor usage. The UDC is a large dataset that contains information of around 1,800,000 users, and has a total of 2,323,233,101 unique rows with 5 attributes each. Table \ref{tbl:att_udc} show a description of the attributes.

@@

\subsection{Selection and transformation}
We used the data pre-processed by Murphy-Hill, that is published on Google BigQuery \cite{SnipesETALASD}. This is an alternative version of the original UDC dataset, that is cleaned and preprocessed, so that the transformation phase is simple and focused on our needs. Due to the magnitude of the dataset we only worked on a fragment of it. We took a random sample of 1,000 users, which is large enough to be representative; in fact, this is larger than the dataset we were able to obtain for the Mylyn study.

@@

After obtaining the data, the next step was to add the attributes required by every event in order to match as closely as possible the format of the Mylyn data; this allowed us to reuse a part of our analysis. We hence added fields corresponding to the type of the event (edition or selection), its duration (time elapsing between this event and the next event, used to determine where interruptions take place and sessions end) and an ID to identify the different sessions present on the data. For the latter, we sorted the data by \textit{userId} and by \textit{datetime} afterwards. This was required because by default the user's data is mixed and we needed it not only chronologically correct but also sorted by users to tag the sessions of every user without interferences.

%With Mylyn we have information about the end time of an event. The UDC dataset doesn't have an attribute to show this information, so we added an interval time in seconds that depends on the next event executed. This is important to identify those events that represent an interruption or the end of a session.

Once the events were classified and sorted, we split the dataset into sessions, defining a session as a period of registered work between interruptions of 8 hours or larger, as we did with Mylyn. We did this splitting the data by users to avoid a conflict between the last event of a user and the first one of the next. After this, we obtained a total of 28,989 different sessions; after a filter to keep only those with at least 30 minutes of productive time (as we did with the Mylyn dataset), we ended up with 15,825 useful sessions, in contrast with the 4,284 we had from Mylyn.

\subsection{Classification of events}
Mylyn events belonged to two principal categories: edition and selection. The UDC data doesn't have this, so we inferred a classification manually. Note that we reuse the same term of selection in the following text, as it was used in Mylyn data. Strictly speaking, the word "navigation" would be better suited to the variety of events present in the UDC data, but we keep the old term for consistency.
%To select the event type, we made a thorough analysis of the different events to classify them where they better fit. In comparison with Mylyn we didn't have a classification by default but the attributes can give us a very good approximation.

We took into consideration the attributes \textit{what}, \textit{kind} and \textit{description} to choose the best classification for every event. Starting by the kind attribute, some events are labeled as \textit{view} which are precisely those that open, activate or close a view such as the console, the package explorer or the variables view. Therefore, the events of the kind \textit{view} were classified as selection-type events.

The classification for the rest of the events is more complex: we needed to check directly the command description and the bundle name it belongs to in order to precisely infer its meaning (e.g., . We could label as edition all those commands with the attributes \textit{command} and \textit{executed}. The majority of them are in fact of that type, but some commands executed also can open a view, navigate through the text or go to another class. For this reason we took a closer look into the bundle name and what the description can tell us about the events.

Thus, by default, all events of the attributes \textit{command} and \textit{executed} were classified as edition and afterwards we did a thorough analysis to identify those that are actually selection events.  This way we categorized as selection all the events where we could tell based on the description if they have the purpose to, for instance, navigate to another window, open a perspective, find and replace words, search code, debug code, or perform window management. The description contains the whole path from the top package to the class that implements the event, so we look for key words to apply the classification. The proportion of events after the classification is shown in the table \ref{tbl:prop_events_udc}.
We used an iterative process involving two of the authors of the paper to arrive to this result; the process was iterative, until we reached consensus.

@@

The edition and selection types the ones we are interested in for the study. The system classification is composed of the events that activate and deactivate the workspace of Eclipse; it is difficult to determine when and why this events are executed (they can be executed either via user or system action, and the order of execution in relation with associated events is not clear) so we decided to establish this particular classification for them. Finally, the events of the type control are those for file version control.
